Measuring individual differences in the understanding of gaze cues across the lifespan

### 2022-04-11 paper/manuscript.tex
edit methods section, add abstract first draft

\abstract{%
There must be an abstract of no more than 250 words.
One or two sentences providing a \textbf{basic introduction} to the field, comprehensible to a scientist in any discipline.

Two to three sentences of \textbf{more detailed background}, comprehensible to scientists in related disciplines.

One sentence clearly stating the \textbf{general problem} being addressed by this particular study.

One sentence summarizing the main result (with the words ``\textbf{here we show}'' or their equivalent).

Two or three sentences explaining what the \textbf{main result} reveals in direct comparison to what was thought to be the case previously, or how the main result adds to previous knowledge.

One or two sentences to put the results into a more \textbf{general context}.

Two or three sentences to provide a \textbf{broader perspective}, readily comprehensible to a scientist in any discipline.
}

@@

Our newly implemented balloon finding task displays cartoon-like stimuli that are engaging for children and adults alike. An animal character (i.e., agent; sheep, monkey, or pig) is placed centrally in a window. A balloon (i.e., target; blue, green, yellow, or red) is located in front of them. The target then falls to the ground. At all times, the agent's gaze tracks the movement of the target. That is, the pupils and iris of the agent move in a way that their center aligns with the center of the target. While the distance of the target's flight depends on the final location, the target moves at a constant speed. Participants are then asked to locate the target's destination: they respond by touching or clicking onto position on the screen.

To keep participants engaged and interested, the presentation of events is accompanied by cartoon-like effects. Each trial starts with an attention-getter: an eye-blinking sound plays while the pupils and iris of the agent enlarge (increase to 130\%) and change in opacity (decrease to 75\%) for 0.3 sec.~The landing of the target is accompanied by a tapping sound. Once the target landed, the instructor's voice asks ``Where is the balloon?''. After the response is registered, a short plop sound plays and a small orange circle confirms the participants' location choice. If no response got registered within 5 secs after the target landed, an audio prompt reminds the participant to respond.

@@

Before the test trials start, we present four training trials during which participants familiarize themselves with selection positions on the screen. In the first training trial, participants have full visual access to the target flight and the target's end location and are simply asked to click on the visible balloon. In the second and third training trials, participants have partial visual access: they witness the target flight but cannot see the target's end location. They are then asked to click on the invisible balloon, i.e., the location where they saw the target land. In test trials, participants have no visual access to the target flight nor the end location. Participants are expected to use the agent's gaze as a cue to locate the target. The first trial of each type comprises a voice-over description of the presented events. The audio descriptions explicitly state that the agent is always looking at the target (see Appendix for audio script). After these four training and audio guided trials, participants receive 15 test trials. The complete sequence of four training trials and 15 test trials can be administered within 5-10 minutes of testing time.

@@

We designed two study versions which can be chosen according to the researchers' need: there is a continuous \emph{hedge} version and a discrete \emph{box} version. Both versions use the same first training trial and then differ in the consecutive training and test trials.
In the \emph{hedge} version, participants have to indicate their estimated target location directly on a hedge (i.e., hedge version).
Here, the dependent variable is imprecision, which is defined as the absolute difference between the target's true x coordinate and the x coordinate of the participant's click.
In the \emph{box} version, participants are asked to click on a box that hides the target. Researchers have the choice of how many boxes are shown: one up to eight boxes can be displayed as potential hiding locations. Here, we use a categorical outcome (i.e., which box was clicked) to calculate the proportion of correct responses.
Note that in the test trials of both study versions, the target flight is covered by a hedge. In the \emph{hedge} version, the hedge then shrinks to a minimum height required to cover the target's end location. In the \emph{box} version, the hedge shrinks completely. The boxes then hide the target's final destination.

@@

\caption{\textbf{Study setup}.
(a) Infrastructure for online testing. (i) Subjects aged 3 -- 99+ can participate. Data collection can take place anywhere: at home, in kindergartens or research institutes. (ii) The task is presented as a website that works across devices. (iii) The scripts for the website and the recorded data are stored on secure local servers.
(b) Continuous \emph{hedge} version of the balloon finding task. (i) The agent stands in a window with the target in front of them. (ii) A hedge grows and covers the target. (iii) The target falls to a random location on the ground. The agent's eyes track the movement of the target.
(c) Discrete \emph{box} version of the balloon finding task. Number of boxes (min. 1; max. 8) as potential hiding locations can be set individually.}\label{fig:fig1}
\end{figure}

@@

All agents and target colors appear equally often and are not repeated in more than two consecutive trials. The randomization of the target end location depends on the study version. In the \emph{hedge} version, the full width of the screen is divided into ten bins. Exact coordinates within each bin are then randomly generated. In the \emph{box} version, the target randomly lands in one of the boxes. As with agent and color choice, each bin/box occurs equally often and can only occur twice in a row.

@@

Our balloon finding task is presented as an interactive web-app. The task is portable across devices and web browsers and does not require any installation. A great advantage of online testing is that our testing procedure is standardized across participants. By using pre-recorded study instructions, no interaction with the experimenter is necessary during the study. The code is open-source (\url{https://github.com/ccp-eva/gafo-demo}) and a live demo version can be found under: \url{https://ccp-odc.eva.mpg.de/gafo-demo/}.
The web-app was programmed in JavaScript (ECMAScript 2015, i.e., ES6), HTML5, CSS and PHP.
We utilized the zero configuration bundler Parcel to enable a live server for debugging, easy imports and minified scripts in order to save data storage.
For stimulus presentation, a scalable vector graphic (SVG) composition was parsed. This way, the composition scales according to the user's view port without loss of quality, while keeping the aspect ratio and relative object positions constant.
Furthermore, SVGs allow us to define all composite parts of the scene (e.g., pupil of the agent) individually. This is needed for precisely calculating exact pupil and target locations and sizes. Additionally, it makes it easy to adjust the stimuli and, for example, add another agent to the scene.
The GreenSock Animation Platform (GSAP; Greensock (2022)) library was used to animate the movement of single SVG elements.
We use URL parameters to capture the participant's ID, language and study version.
The web-app generates two file types: (1) a text file (.json) containing meta-data, trial specifications and participants' click responses, and (2) a video file (.webm) of the participant's webcam recording.
For our samples described in this paper, we deployed the web-app on servers located in Leipzig, Germany. Data got automatically collected and safely stored on these in-house servers. If no internet connection or server is available, researchers could download a local version of the experiment which stores the generated data automatically on the used device. For child samples, we upload safety copies of the already collected responses after the fourth test trial. In cases where children want to stop participation earlier, no responses get lost.

@@

For our remote child sample, we recruited families on a voluntary basis via email from the institute's internal database. Our remote child sample included 147 children, including
45 3-year-olds
(mean = 42.62 months,
SD = 3.35,
range = 36
- 47,
14 girls),
47 4-year-olds
(mean = 52.64 months,
SD = 3.40,
range = 48
- 59,
25 girls),
and 55 5-year-olds
(mean = 65.11 months,
SD = 3.77,
range = 60
- 71,
27 girls). Children in our sample grow up in an industrialized, urban Central-European context. Information on socioeconomic status was not formally recorded, although the majority of families come from mixed, mainly mid to high socioeconomic backgrounds with high levels of parental education.

In addition, we recruited a remote adult sample by advertising the study on \emph{Prolific}. \emph{Prolific} is an online participant recruitment service from the University of Oxford with a predominantly European and US-american subject pool. Participants consisted of 100 English-speakers with an average age of 31.34 years (SD = 10.77, range = 18 - 63, 64 females).
For completing the study, subjects were payed above the fixed minimum wage (in average £10.00 per hour). \emph{Prolific} distributed our study link to potential participants, while the hosting of the online study was done by local servers in the Max Planck Institute for Evolutionary Anthropology, Leipzig, Germany.

@@

Children in our in-person sample were tested on a tablet in a quiet room in their daycare center. An experimenter accompanied the child and helped them navigate through the online study.
Children in the remote sample received a personalized link to the study websites and families could participate at any time or location they wanted. In the beginning of the online study, families were invited to enter our ``virtual institute'' and were welcomed by an introductory video of the study leader, shortly describing the research background and further procedure. Then, caregivers were informed about data security and were asked for their informed consent. They were asked to enable the sound and seat their child centrally in front of their device. Before the study started, families were instructed how to setup their webcam and enable the recording permissions. We stressed that caregivers should not help their children. Study participation was video recorded whenever possible in order to ensure that the answers were generated by the children themselves.
Depending on the participant's device, the website automatically presented the hedge or box version of the study. For families that used a tablet with touchscreen, the hedge version was shown. Here, children could directly click on the touchscreen themselves to indicate where the target is. For families that used a computer without touchscreen, the website presented the box version of the task. We assumed that younger children in our sample would not be acquainted with the usage of a computer mouse. Therefore, we asked children to point to the screen, while caregivers were asked to act as the ``digital finger'' of their children and click on the indicated box.

@@

To estimate the developmental trajectory of gaze cue understanding and the effect of data collection mode, we fit a GLMM predicting the task performance by age (in months, z-transformed) and data collection mode (reference category: in-person supervised). The model included random intercepts for each participant and each target position, and a random slope for symmetric target position within participants (model notation in \texttt{R:\ performance\ \textasciitilde{}\ age\ +\ datacollection\ +\ (symmetricPosition\ \textbar{}\ subjID)\ +\ (1\ \textbar{}\ targetPosition)}). Here, \texttt{targetPosition} refers to the absolute bin/box of the target, while \texttt{symmetricPosition} refers to the distance from the stimulus center (i.e., smaller value meaning more central target position). We expected that trials could differ in their difficulty depending on the target centrality and that these these item effects could vary between participants.

@@

\caption{\textbf{Measuring interindividual variation}.
(a) Developmental trajectory in continuous \emph{hedge} version. Performance is measured as average imprecision, i.e., the absolute distance between the target's center and the participant's click. The unit of imprecision is counted in the width of the target, i.e., a participant with an imprecision of 1 clicked in average one target width to the left or right of the true target center.
(b) Internal consistency (odd-even split) in hedge child sample. (c) Internal consistency in hedge adult sample. (d) Test-retest reliability in hedge child sample. (e) Test-retest reliability in hedge adult sample.
(f) Developmental trajectory in discrete \emph{box} version. Performance is measured as the proportion of correct responses, i.e., how many times the participant clicked on the box that actually contained the target. Dotted black line shows level of performance expected by chance (for child sample 20\%, i.e., 1 out of 5 boxes; for adult sample 12.5\%, i.e., 1 out of 8 boxes).
(g) Internal consistency (odd-even split) in box child sample. (h) Internal consistency in box adult sample. (i) Test-retest reliability in box child sample. (j) Test-retest reliability in box adult sample.
Regression lines with 95\% CI show product-moment-correlations with \emph{Pearson}'s correlation coefficient \emph{r.}
Large data points with 95\% CI (based on non-parametric bootstrap) represent performance means by age group (binned by year).
Small data points show the mean performance for each subject. Shape of data points represents data collection mode: opaque round circles stand for in-person supervised data collection, translucent diamonds stand for remote unsupervised data collection. Color of data points represent each age group (binned by year).}\label{fig:fig2}
\end{figure}

@@

Study 1 showed that our newly implemented balloon finding task can be used to study gaze cue understanding in both children and adults. With increasing age, participants got more and more precise in locating the target. We found interindividual variation across all age groups. Furthermore, we found a comparable developmental trajectory for an unsupervised remote child sample. This underlines how flexibly our new task can be used.

@@

For our child sample, we went to kindergartens in Leipzig and surroundings.
The child sample consisted of
93 children, including
24 3-year-olds
(mean = 42.49 months,
SD = 2.99,
range = 38
- 47,
13 girls),
38 4-year-olds
(mean = 53.77 months,
SD = 3.16,
range = 48
- 59,
19 girls),
and 31 5-year-olds
(mean = 66 months,
SD = 3.42,
range = 61
- 71,
17 girls).

The adult sample was recruited over \emph{Prolific} and consisted of 136 English-speakers with an average age of 25.74 years (SD = 8.11, range = 18 - 71, 87 females).

@@

We assessed reliability in two ways.
First, we focused on the internal consistency by calculating splithalf reliability coefficients. For each subject, trials were split into odd and even and performance scores per part were correlated against each other using \emph{Pearson} correlation coefficients. We took the data of the first test day and defined performance according to study version: in the hedge version, performance referred to the absolute difference between the target center and the click coordinate, scaled according to target widths; in the box version, we computed the proportion of correct choices.
As a recent paper by Pronk, Molenaar, Wiers, and Murre (2021) suggested, there are various less well-known split-half methods that differ in how the trials are split into parts and whether they are combined with stratification by task design. To compare our traditional approach of a simple odd-even split, we additionally calculated reliability estimates using first-second, odd-even, permutated, and Monte Carlo splits without and with stratification by target position. First-second and odd-even splits belong to single sample methods, since each participant has a single pair of performance scores, while permutated (without replacement) and Monte Carlo (with replacement) splits make use of resampling. Analyses were run using the function \texttt{by\_split}from the \texttt{splithalfr} package (Pronk et al., 2021).
Second, we assessed the test-retest reliability. We calculated performance scores (depending on study version as described above) for each individual in each test session and correlated those using \emph{Pearson} correlation coefficients. Furthermore, for our child sample we report an age-corrected correlation between the two test days using a GLMM based approach. We fit trial by trial data with a fixed effect of age, a random intercept for each subject and a random slope for test day (model notation in \texttt{R:\ performance\ \textasciitilde{}\ age\ (0\ +\ reliday\ \textbar{}\ subjID)}). For the hedge version, performance was modeled by a lognormal distribution, while the model for the box version used a Bernoulli distribution with a logit link function.
The model computes a correlation between the participant specific estimates for each test day. This can be interpreted as the test-retest reliability. By using this approach, we do not need to compromise on data aggregation and, therefore, loss of information. Most importantly, the model allows us to get an age-independent estimate for reliability. This rules out the possibility that a high correlation between test days arises from domain general cognitive maturation processes instead of study-specific inter-individual differences.

@@

For this exploratory analysis, we included all children of the aforementioned samples where families filled out a short demographic questionnaire. This subsample consisted of
130 children, including
39 3-year-olds
(mean = 43.02 months,
SD = 3.20,
range = 37
- 47,
20 girls),
44 4-year-olds
(mean = 54.43 months,
SD = 2.77,
range = 48
- 59,
26 girls),
and 47 5-year-olds
(mean = 66.13 months,
SD = 3.55,
range = 60
- 71,
23 girls).

@@

To estimate the effects of social surrounding on gaze cue understanding, we fit GLMMs predicting the task performance by each of our questionnaire variables, controlling for age (in months, z-transformed), data collection mode (reference category: in-person supervised) and study version (reference category: hedge version). The models included random intercepts for each participant and each target position, and a random slope for symmetric target position within participants. Therefore, our null model closely resembled the structure from our first analysis (section \emph{Does the balloon finding task induce variation?}; here: \texttt{performance\ \textasciitilde{}\ age\ +\ datacollection\ +\ studyversion\ +\ (symmetricPosition\ \textbar{}\ subjID)\ +\ (1\ \textbar{}\ targetPosition)}).
In order to combine data of our two study versions, we transformed continuous click responses from the hedge version into a discrete outcome. For the target position, we categorized two adjacent bins as one imaginary box. To measure participants' performance, we created imaginary box boundaries around the target's landing position and examined whether the participant's click response fell into this imaginary box. Across the two study versions, we could consequently model the participant's correct response (0/1) using a Bernoulli distribution with a logit link function.
For model comparisons, we respectively added the following predictors as fixed effects to the null model: number of household members, number of children aged 0-18 in household, number of children aged 1-12 in household, hours spent in childcare each day, and age when subject entered childcare.
In addition, we calculated three index scores. First, we calculated a sibling variety score according to Peterson (2000). Second, we implemented the modified version of Cassidy, Fineberg, Brown, and Perkins (2005) (for more details, see Supplements). Third, based on our own data exploration, we calculated the amount of peer exposure determined as the number of siblings and the average hours spent in childcare (both z-transformed).
We compared the models using WAIC (widely applicable information criterion) scores and weights. As an indicator of out-of-sample predictive accuracy, lower WAIC scores stand for a better model fit. WAIC weights represent the probability that the model in question provides the best out-of-sample prediction compared to the other models.

@@

The model including our peer exposure index, as defined as the number of other children in the household and average hours spent in childcare, showed the best out-of-sample predictive accuracy. Note that we did not find a great difference in WAIC scores between the compared models (see Supplements for WAIC scores and weights). The model estimates were all considerably smaller than estimates of age, study version and data collection, and all 95\% CIs included zero. For example, for our winning model, we found an estimate of
0.18 (95\% CI {[}-0.01; 0.36{]}),
with the estimates of age being 0.59 (95\% CI {[}0.40; 0.79{]}), data collection mode being 0.93 (95\% CI {[}0.54; 1.34{]}), and study version 1.84 (95\% CI {[}0.14; 3.47{]}). Nevertheless, a general direction of effect can be seen: there seems to be small evidence that a social surrounding positively influenced children's gaze cue understanding. The number of people and, more specifically, children, as well as the more diverse their age, the more likely children were to follow the agent's gaze. The only predictor resulting in a negative estimate was the age at which a participant entered childcare, i.e., the later a child entered, the more likely children found the target.

@@

\caption{\textbf{External validity of the balloon finding task}.
Factors of children's social surroundings and their influence on the probability of responding correctly. Models are ordered according to their WAIC scores, with the uppermost winning the model comparison. The graph shows the estimated density curves of a model's predictor coefficients. Red tails display the 2.5\% and 97.5\% quantiles, respectively.}\label{fig:fig4}
\end{figure}

>>>>

\abstract{%
In order to explain and predict the behavior of agents, we use social cognition: we represent and reason about other's perspectives, knowledge, intentions, beliefs and preferences. Traditional measures of social cognition (e.g., false belief change-of-location tasks), however, often lack satisfactory psychometric properties: they are not designed to capture variation \emph{between} children and rely on low trial numbers, dichotomous measures, and group averages. This has profound implications on what these studies can show. Poor measurement of social cognition on an individual level may conceal relations between different aspects of cognition and may obscure developmental change. To fully understand how social-cognitive abilities emerge and relate to each other, we need new tools that can reliably measure individual differences.
To approach this issue, we designed a balloon finding task to study social cognition in young children and adults. We concentrate on an essential ability that is involved in many social-cognitive reasoning processes: gaze cue understanding -- the ability to locate and use the attentional focus of an agent. Our interactive web interface works across devices and enables supervised and unsupervised, as well as in-person and remote testing. The implemented spatial layout allows for discrete as well as continuous measures of participants' click imprecision and is easily adoptable to different study needs.
Here we show that our task induces inter-individual differences in a child (N = XXX) and adult (N = XXX) sample. Our two study versions and data collection modes yield comparable results that show substantial developmental gains: the older children are, the more accurately they locate the target. High internal consistency and test-test reliability estimates underline that the capured variation is systematic. Furthermore, we find first evidence for the external validity of our task: the measured performance in gaze cue understanding relates to children's real life social surrounding. Taken together, this work shows a promising way forward in the study of individual differences in social cognition and will help us explore the in(ter)dependence of our core social-cognitive processes in greater detail.
}

@@

Our newly implemented task features an online game where children or adults are asked to search for a balloon. The events proceed as follows. An animated agent (a sheep, monkey, or pig) looks out of a window of a house. A balloon (i.e., target; blue, green, yellow, or red) is located in front of them. The target then falls to the ground. At all times, the agent's gaze tracks the movement of the target. That is, the pupils and iris of the agent move in a way that their center aligns with the center of the target. While the distance of the target's flight depends on the final location, the target moves at a constant speed. Participants are then asked to locate the target's destination: they respond by touching or clicking onto a position on the screen. Visual access to the target's true location is manipulated by a hedge. Participants either have full, partial, or no visual access to the true target location. When partial or no information about the target location is accessible, participants are expected to use the agent's gaze as a cue.

To keep participants engaged and interested, the presentation of events is accompanied by cartoon-like effects. Each trial starts with an attention-getter: an eye-blinking sound plays while the pupils and iris of the agent enlarge (increase to 130\%) and change in opacity (decrease to 75\%) for 0.3 sec.~The landing of the target is accompanied by a tapping sound. Once the target landed, the instructor's voice asks ``Where is the balloon?''. For confirming the participant's click, a short plop sound plays and a small orange circle appears at the location of choice. If no response got registered within 5 secs after the target landed, an audio prompt reminds the participant to respond.

@@

Trials differ in the amount of visual access that participants have to the final target position. Before the test trials start, we present four training trials during which participants familiarize themselves with selecting positions on the screen. In the first training trial, participants have full visual access to the target flight and the target's end location and are simply asked to click on the visible balloon. In the second and third training trials, participants have partial visual access: they witness the target flight but cannot see the target's end location. They are then asked to click on the invisible balloon, i.e., the location where they saw the target land. In test trials, participants have no visual access to the target flight nor the end location. Participants are expected to use the agent's gaze as a cue to locate the target. The first trial of each type comprises a voice-over description of the presented events. The audio descriptions explicitly state that the agent is always looking at the target (see Appendix for audio script). After these four training and audio guided trials, participants receive 15 test trials. The complete sequence of four training trials and 15 test trials can be administered within 5-10 minutes of testing time.

@@

We designed two study versions which can be chosen according to the researchers' need. The study versions differ in the final hiding place of the target and, consequently, on the outcome measure: there is a continuous hedge version and a discrete box version. Both versions use the same first training trial and then differ in the consecutive training and test trials.
In the hedge version, participants have to indicate their estimated target location directly on a hedge (i.e., hedge version).
Here, the dependent variable is imprecision, which is defined as the absolute difference between the target center and the x coordinate of the participant's click.
In the discrete study version, the target lands in a box and participants are asked to click on the box that hides the target. Researchers have the choice how many boxes are shown: one up to eight boxes can be displayed as potential hiding locations. Here, we use a categorical outcome (i.e., which box was clicked) to calculate the proportion of correct responses.
Note that in the test trials of both study versions, the target flight is covered by a hedge. In the hedge version, the hedge then shrinks to a minimum height required to cover the target's end location. In the box version, the hedge shrinks completely. The boxes then hide the target's final destination.

@@

\caption{\textbf{Study setup}.
(a) Infrastructure for online testing. (i) Subjects aged 3 -- 99+ can participate. Data collection can take place anywhere: at home, in kindergartens or research institutes. (ii) The task is presented as a website that works across devices. (iii) The scripts for the website and the recorded data are stored on secure local servers.
(b) Continuous hedge version of the balloon finding task. (i) The agent stands in a window with the target in front of them. (ii) A hedge grows and covers the target. (iii) The target falls to a random location on the ground. The agent's eyes track the movement of the target.
(c) Discrete box version of the balloon finding task. Number of boxes (min. 1; max. 8) as potential hiding locations can be set according to the researcher's need.}\label{fig:fig1}
\end{figure}

@@

All agents and target colors appear equally often and are not repeated in more than two consecutive trials. The randomization of the target end location depends on the study version. In the hedge version, the full width of the screen is divided into ten bins. Exact coordinates within each bin are then randomly generated. In the box version, the target randomly lands in one of the boxes. As with agent and color choice, each bin/box occurs equally often and can only occur twice in a row.

@@

Our balloon finding task is presented as an interactive web-app. The task is portable across devices and web browsers and does not require any installation. A great advantage of online testing is that our testing procedure is standardized across participants. By using pre-recorded study instructions, no interaction with the experimenter is necessary during the study. The code is open-source (\url{https://github.com/ccp-eva/gafo-demo}) and a live demo version can be found under: \url{https://ccp-odc.eva.mpg.de/gafo-demo/}.
The web-app was programmed in JavaScript (ECMAScript 2015, i.e., ES6), HTML5, CSS and PHP.
We utilized the zero configuration bundler Parcel to enable a live server for debugging, easy imports and minified scripts in order to save data storage.
For stimulus presentation, a scalable vector graphic (SVG) composition was parsed. This way, the composition scales according to the user's view port without loss of quality, while keeping the aspect ratio and relative object positions constant.
Furthermore, SVGs allow us to define all composite parts of the scene (e.g., pupil of the agent) individually. This is needed for precisely calculating exact pupil and target locations and sizes. Additionally, it makes it easy to adjust the stimuli and, for example, add another agent to the scene.
The GreenSock Animation Platform {[}GSAP; Greensock (2022){]} library was used to animate the movement of single SVG elements.
We use URL parameters to capture the participant's ID, language and study version.
The web-app generates two file types: (1) a text file (.json) containing meta-data, trial specifications and participants' click responses, and (2) a video file (.webm) of the participant's webcam recording.
For our samples described in this paper, we deployed the web-app on servers located in Leipzig, Germany. Data got automatically collected and safely stored on these in-house servers. If no internet connection or server is available, researchers could download a local version of the experiment which stores the generated data automatically on the used device. For child samples, we upload safety copies of the already collected responses after the fourth test trial. In cases where children want to stop participation earlier, no responses get lost.

@@

For our remote child sample, we recruited families on a voluntary basis via email from our institute's internal database. The remote child sample included 147 children, including
45 3-year-olds
(mean = 42.62 months,
SD = 3.35,
range = 36
- 47,
14 girls),
47 4-year-olds
(mean = 52.64 months,
SD = 3.40,
range = 48
- 59,
25 girls),
and 55 5-year-olds
(mean = 65.11 months,
SD = 3.77,
range = 60
- 71,
27 girls). Children in our sample grow up in an industrialized, urban Central-European context. Information on socioeconomic status was not formally recorded, although the majority of families come from mixed, mainly mid to high socioeconomic backgrounds with high levels of parental education.

In addition, we recruited a remote adult sample by advertising the study on \emph{Prolific}. \emph{Prolific} is an online participant recruitment service from the University of Oxford with a predominantly European and US-american subject pool. Participants consisted of 50, 50, 50, and 50 English-speakers with an average age of 31.92, 31.92, 30.76, and 30.76 years (SD = 12.15, 12.15, 9.12, and 9.12, range = 18, 18, 19, and 19 - 63, 63, 59, and 59, 36, 28, 36, and 28 females).
For completing the study, subjects were payed above the fixed minimum wage (in average £10.00 per hour). \emph{Prolific} distributed our study link to potential participants, while the hosting of the online study was done by local servers in the Max Planck Institute for Evolutionary Anthropology, Leipzig, Germany.

@@

Children in our in-person sample were tested on a tablet in a quiet room in their daycare center. An experimenter accompanied the child and helped them navigate through the online study.

Children in the remote sample received a personalized link to the study website and families could participate at any time or location they wanted. In the beginning of the online study, families were invited to enter our ``virtual institute'' and were welcomed by an introductory video of the study leader, shortly describing the research background and further procedure. Then, caregivers were informed about data security and were asked for their informed consent. They were asked to enable the sound and seat their child centrally in front of their device. Before the study started, families were instructed how to setup their webcam and enable the recording permissions. We stressed that caregivers should not help their children. Study participation was video recorded whenever possible in order to ensure that the answers were generated by the children themselves.
Depending on the participant's device, the website automatically presented the hedge or box version of the study. For families that used a tablet with touchscreen, the hedge version was shown. Here, children could directly click on the touchscreen themselves to indicate where the target is. For families that used a computer without touchscreen, the website presented the box version of the task. We assumed that younger children in our sample would not be acquainted with the usage of a computer mouse. Therefore, we asked children to point to the screen, while caregivers were asked to act as the ``digital finger'' of their children and click on the indicated box.

@@

To estimate the developmental trajectory of gaze cue understanding and the effect of data collection mode, we fit a GLMM predicting the task performance by age (in months, z-transformed) and data collection mode (reference category: in-person supervised). The model included random intercepts for each participant and each target position, and a random slope for symmetric target position within participants (model notation in \texttt{R:\ performance\ \textasciitilde{}\ age\ +\ datacollection\ +\ (symmetricPosition\ \textbar{}\ subjID)\ +\ (1\ \textbar{}\ targetPosition)}). Here, \texttt{targetPosition} refers to the exact bin/box of the target, while \texttt{symmetricPosition} refers to the absolute distance from the stimulus center (i.e., smaller value meaning more central target position). We expected that trials could differ in their difficulty depending on the target centrality and that these these item effects could vary between participants.

@@

\caption{\textbf{Measuring interindividual variation}.
(a) Developmental trajectory in continuous hedge version. Performance is measured as average imprecision, i.e., the absolute distance between the target's center and the participant's click. The unit of imprecision is counted in the width of the target, i.e., a participant with an imprecision of 1 clicked in average one target width to the left or right of the true target center.
(b) Internal consistency (odd-even split) in hedge child sample. (c) Internal consistency in hedge adult sample. (d) Test-retest reliability in hedge child sample. (e) Test-retest reliability in hedge adult sample.
(f) Developmental trajectory in discrete box version. Performance is measured as the proportion of correct responses, i.e., how many times the participant clicked on the box that actually contained the target. Dotted black line shows level of performance expected by chance (for child sample 20\%, i.e., 1 out of 5 boxes; for adult sample 12.5\%, i.e., 1 out of 8 boxes).
(g) Internal consistency (odd-even split) in box child sample. (h) Internal consistency in box adult sample. (i) Test-retest reliability in box child sample. (j) Test-retest reliability in box adult sample.
Regression lines with 95\% CI show product-moment-correlations with \emph{Pearson}'s correlation coefficient \emph{r.}
Large data points with 95\% CI (based on non-parametric bootstrap) represent performance means by age group (binned by year).
Small data points show the mean performance for each subject. Shape of data points represents data collection mode: opaque round circles stand for in-person supervised data collection, translucent diamonds stand for remote unsupervised data collection. Color of data points represent each age group (binned by year).}\label{fig:fig2}
\end{figure}

@@

Our newly implemented balloon finding task can be used to study gaze cue understanding in both children and adults. With increasing age, participants got more and more precise in locating the target. We found interindividual variation across all age groups. Furthermore, we found a comparable developmental trajectory for an unsupervised remote child sample. This underlines how flexibly our new task can be used.

@@

For our child sample, we went to kindergartens in Leipzig and surroundings.
The child sample consisted of
106 children, including
35 3-year-olds
(mean = 42.57 months,
SD = 2.98,
range = 38
- 47,
17 girls),
38 4-year-olds
(mean = 53.77 months,
SD = 3.16,
range = 48
- 59,
19 girls),
and 33 5-year-olds
(mean = 66.12 months,
SD = 3.36,
range = 61
- 71,
17 girls).

The adult sample was recruited over \emph{Prolific} and consisted of 70, 70, 66, and 66 English-speakers with an average age of 25.43, 25.43, 26.05, and 26.05 years (SD = 6.43, 6.43, 9.44, and 9.44, range = 18, 18, 18, and 18 - 51, 51, 71, and 71, 45, 42, 45, and 42 females).

@@

We assessed reliability in two ways.
First, we focused on the internal consistency by calculating splithalf reliability coefficients. For each subject, trials were split into odd and even and performance scores per part were correlated against each other using \emph{Pearson} correlation coefficients. We took the data of the first test day and defined performance according to study version: in the hedge version, performance referred to the absolute difference between the target center and the click coordinate, scaled according to target widths; in the box version, we computed the proportion of correct choices.
As a recent paper by Pronk, Molenaar, Wiers, and Murre (2021) suggested, there are various less well-known split-half methods that differ in how the trials are split into parts and whether they are combined with stratification by task design. To compare our traditional approach of a simple odd-even split, we additionally calculated reliability estimates using first-second, odd-even, permutated, and Monte Carlo splits without and with stratification by target position. First-second and odd-even splits belong to single sample methods, since each participant has a single pair of performance scores, while permutated (without replacement) and Monte Carlo (with replacement) splits make use of resampling. Analyses were run using the function \texttt{by\_split} from the \texttt{splithalfr} package (Pronk et al., 2021).
Second, we assessed the test-retest reliability. We calculated performance scores (depending on study version as described above) for each individual in each test session and correlated those using \emph{Pearson} correlation coefficients. Furthermore, for our child sample we report an age-corrected correlation between the two test days using a GLMM based approach. We fit trial by trial data with a fixed effect of age, a random intercept for each subject and a random slope for test day (model notation in \texttt{R:\ performance\ \textasciitilde{}\ age\ (0\ +\ reliday\ \textbar{}\ subjID)}). For the hedge version, performance was modeled by a lognormal distribution, while the model for the box version used a Bernoulli distribution with a logit link function.
The model computes a correlation between the participant specific estimates for each test day. This can be interpreted as the test-retest reliability. By using this approach, we do not need to compromise on data aggregation and, therefore, loss of information. Most importantly, the model allows us to get an age-independent estimate for reliability. This rules out the possibility that a high correlation between test days arises from domain general cognitive maturation processes instead of study-specific inter-individual differences.

@@

For this exploratory analysis, we included all children of the aforementioned samples where families filled out a short demographic questionnaire. This subsample consisted of
136 children, including
41 3-year-olds
(mean = 43.17 months,
SD = 3.13,
range = 37
- 47,
20 girls),
46 4-year-olds
(mean = 54.43 months,
SD = 2.76,
range = 48
- 59,
26 girls),
and 49 5-year-olds
(mean = 66.25 months,
SD = 3.47,
range = 60
- 71,
23 girls).

@@

To estimate the effects of social surrounding on gaze cue understanding, we fit GLMMs predicting the task performance by each of our questionnaire variables, controlling for age (in months, z-transformed), data collection mode (reference category: in-person supervised) and study version (reference category: hedge version). The models included random intercepts for each participant and each target position, and a random slope for symmetric target position within participants. Therefore, our null model closely resembled the structure from our first analysis (see Analysis section of \emph{Does the balloon finding task induce variation?}; here: \texttt{performance\ \textasciitilde{}\ age\ +\ datacollection\ +\ studyversion\ +\ (symmetricPosition\ \textbar{}\ subjID)\ +\ (1\ \textbar{}\ targetPosition)}).
In order to combine data of our two study versions, we transformed continuous click responses from the hedge version into a discrete outcome. For the target position, we categorized two adjacent bins as one imaginary box. To measure participants' performance, we created imaginary box boundaries around the target's landing position and examined whether the participant's click response fell into this imaginary box. Across the two study versions, we could consequently model the participant's correct response (0/1) using a Bernoulli distribution with a logit link function.
For model comparisons, we respectively added the following predictors as fixed effects to the null model: number of household members, number of children aged 0-18 in household, number of children aged 1-12 in household, hours spent in childcare each day, and age when subject entered childcare.
In addition, we calculated three index scores. First, we calculated a sibling variety score according to Peterson (2000). Second, we implemented the modified version of Cassidy, Fineberg, Brown, and Perkins (2005) (for more details, see Supplements). Third, based on our own data exploration, we calculated the amount of peer exposure determined as the number of siblings and the average hours spent in childcare (both z-transformed).
We compared the models using WAIC (widely applicable information criterion) scores and weights. As an indicator of out-of-sample predictive accuracy, lower WAIC scores stand for a better model fit. WAIC weights represent the probability that the model in question provides the best out-of-sample prediction compared to the other models.

@@

The model including our peer exposure index, as defined as the number of other children in the household and average hours spent in childcare, showed the best out-of-sample predictive accuracy. Note that we did not find a great difference in WAIC scores between the compared models (see Supplements for WAIC scores and weights). The model estimates were all considerably smaller than estimates of age, study version and data collection, and all 95\% CIs included zero. For example, for our winning model, we found a peer exposure estimate of
0.18 (95\% CI {[}-0.01; 0.36{]}),
with the estimates of age being 0.59 (95\% CI {[}0.40; 0.79{]}), data collection mode being 0.93 (95\% CI {[}0.54; 1.34{]}), and study version 1.84 (95\% CI {[}0.14; 3.47{]}). Nevertheless, a general direction of effect can be seen: there seems to be small evidence that a social surrounding positively influenced children's gaze cue understanding. The number of people and, more specifically, children, as well as the more diverse their age, the more likely children were to follow the agent's gaze. The only predictor resulting in a negative estimate was the age at which a participant entered childcare, i.e., the later a child entered, the more likely children found the target.

@@

\caption{\textbf{External validity of the balloon finding task}.
Factors of children's social surroundings and their influence on the probability of responding correctly. Models are ordered according to their WAIC scores, with the uppermost winning the model comparison. The graph shows the estimated density curves of a model's predictor coefficient. Red tails display the 2.5\% and 97.5\% quantiles, respectively.}\label{fig:fig4}
\end{figure}
